The Early Arrival of WASP-4\lowercase{b}

### 2018-12-07 paper/ms.tex
draft: v0.1. ready for JNW.

%\documentclass[preprint2,tighten]{aastex6}
\documentclass[12pt,twocolumn,tighten]{aastex62}
%\pdfoutput=1 %for arXiv submission
\usepackage{amsmath,amstext,amssymb}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage[figure,figure*]{hypcap}
\usepackage{graphics,graphicx}
\usepackage{hyperref}
\usepackage{comment}
\usepackage{todonotes}

@@

\shortauthors{Bouma et al.}
\shorttitle{Decreasing Period of WASP-4\lowercase{b}}

@@

\title{The Decreasing Orbital Period of WASP-4\lowercase{b}}

@@

\author[0000-0002-0514-5538]{L. G. Bouma}
\affiliation{ Department of Astrophysical Sciences, Princeton
University, 4 Ivy Lane, Princeton, NJ 08540, USA}

\author[0000-0002-4265-047X]{J. N. Winn}
\affiliation{ Department of Astrophysical Sciences, Princeton
University, 4 Ivy Lane, Princeton, NJ 08540, USA}

\author[0000-0002-6939-9211]{T. Daylan}
\affiliation{ Department of Physics and Kavli Institute for Astrophysics
and Space Research, Massachusetts Institute of Technology, Cambridge, MA
02139, USA }

\author[0000-0002-3481-9052]{K. G. Stassun}
\affiliation{Vanderbilt University, Department of Physics \& Astronomy,
6301 Stevenson Center Lane, Nashville, TN 37235, USA}
\affiliation{Fisk University, Department of Physics, 1000 17th Avenue
N., Nashville, TN 37208, USA}

\author[0000-0002-7084-0529]{S. Kane}
\affiliation{Department of Earth Sciences, University of California,
Riverside, CA 92521, USA}

%-------------------------------------
% TESS Mission Architects:
% These authors should be listed in this order
% see https://spacebook.mit.edu/pages/viewpage.action?pageId=24543276
%-------------------------------------
%\\ \textcolor{red}{6 TESS Mission Architects:}
\author{G. R. Ricker} % grr@space.mit.edu WAITING
\affiliation{ Department of Physics and Kavli Institute for Astrophysics
and Space Research, Massachusetts Institute of Technology, Cambridge, MA
02139, USA }
\author[0000-0001-6763-6562]{R. Vanderspek} % roland@space.mit.edu WAITING
\affiliation{ Department of Physics and Kavli Institute for Astrophysics
and Space Research, Massachusetts Institute of Technology, Cambridge, MA
02139, USA }
\author[0000-0001-9911-7388]{D. W.~Latham} % dlatham@cfa.harvard.edu WAITING
\affiliation{ Harvard-Smithsonian Center for Astrophysics, 60 Garden
Street, Cambridge, MA 02138, USA }
\author{S. Seager} % seager@mit.edu WAITING
\affiliation{ Department of Earth, Atmospheric, and Planetary Sciences,
Massachusetts Institute of Technology, Cambridge, MA 02139, USA }
\author[0000-0002-4715-9460]{J. M.~Jenkins} % jon.jenkins@nasa.gov WAITING
\affiliation{ NASA Ames Research Center, Moffett Field, CA 94035, USA }
%-------------------------------------
% 3 representatives of each of SPOC, POC, TSO, for a total of 9. 
%These 9 authors should be listed in alphabetical order
%-------------------------------------
% 3 TSO COAUTHORS
%
\author[0000-0002-3321-4924]{Z. Berta-Thompson} % Zachory.BertaThompson@Colorado.EDU WAITING
\affiliation{Department of Astrophysical and Planetary Sciences,
University of Colorado, Boulder, CO 80309, USA}
%
\author[0000-0001-8812-0565]{J. Rodriguez} % rodriguez.jr.joey@gmail.com WAITING
\affiliation{Harvard-Smithsonian Center for Astrophysics, 60 Garden
Street, Cambridge, MA 02138, USA }
%
\author[0000-0001-8020-7121]{K. Col\'on} % knicole.colon@nasa.gov WAITING
\affiliation{NASA Goddard Space Flight Center, Exoplanets and Stellar
Astrophysics Laboratory (Code 667), Greenbelt, MD 20771, USA}
% 3 SPOC COAUTHORS
%
\author[0000-0002-6778-7552]{J. D. Twicken} % Joseph.Twicken@nasa.gov WAITING
\affiliation{ NASA Ames Research Center, Moffett Field, CA 94035, USA }
\affiliation{ SETI Institute, Mountain View, CA, 94043}
%
\author{B. Wohler} % bill.wohler@nasa.gov WAITING
\affiliation{ NASA Ames Research Center, Moffett Field, CA 94035, USA }
\affiliation{ SETI Institute, Mountain View, CA, 94043}
%
\author[0000-0003-1963-9616]{D. A. Caldwell} % douglas.Caldwell@nasa.gov % WAITING
\affiliation{ NASA Ames Research Center, Moffett Field, CA 94035, USA }
\affiliation{ SETI Institute, Mountain View, CA, 94043}
%
%\\ \textcolor{red}{3 POC coauthors:}
\author{POC1}
\author{POC2}
\author{POC3}

\begin{abstract}
  %TODO: final numbers
  Combining data from the Transiting Exoplanet Survey Satellite (TESS)
  with previous studies, we show that the transit times of WASP-4b are
  incompatible with a constant orbital period. 
  In particular, the transits seem to have arrived $77.8 \pm 10.7$
  seconds early, and the period appears to be shrinking by
  $\dot{P}=-12.1 \pm 1.2$ milliseconds per year.   
  From TESS observations of WASPs-6b, -18b, and -46b, we show
  that a systematic offset between the TESS time stamps and the barycentric
  reference sufficient to explain WASP-4b is ruled out at 6.3$\sigma$.
  If the timing variations are astrophysical,
  major contributors to the period change could include both
  apsidal precession and tidal decay.
  The latter seems implausible on theoretical grounds, as it would
  require a stellar quality factor $Q_\star' \approx 3\times10^4$.
  The Doppler shift from WASP-4's acceleration towards us could
  account for at most one quarter of the observed period change
  (at $2\sigma$).
  Further transit and occultation studies will help confirm the
  reality of the timing variation, and eventually determine its cause.
\end{abstract}

@@

Long-term monitoring of hot Jupiter transit and occultation times should
eventually reveal two distinct processes.  First, whenever the summed
spin and orbital angular momenta in a hot Jupiter system are below a
critical value, the system will be unstable to tidal decay
\citep{counselman_outcomes_1973,hut_stability_1980}.  Most hot Jupiters
satisfy this condition; their orbits should be shrinking
\citep{levrard_falling_2009,matsumura_tidal_2010}.  The rate of decay is
uncertain, and depends on how friction dissipates energy carried by
gravitational tides (\citealt{ogilvie_tidal_2014} gives a review).  If
we could directly measure orbital decay rates, it would inform our
understanding of tidal dissipation.  This is important because the
dissipation rates determine the fate of the system.  During its
inspiral, if the planet fills its Roche lobe below the stellar surface,
the drag force from the stellar envelope accelerates the inspiral,
leading to a `direct-impact' merger
\citep{metzger_optical_2012,macleod_planetary_2018} Alternatively, if
the planet fills its Roche lobe above above the stellar surface, then it
could either be disrupted into an accretion disk or it could
slowly transfer mass to the star
\citep{metzger_optical_2012,valsecchi_tidally-driven_2015,jackson_tidal_2016}.
Further afield, the outcome of the eventual interaction between the
Earth and red giant Sun also hinges on the efficiency of tidal
dissipation \citep{rasio_tidal_1996}.

@@

This work focuses on WASP-4b, which is also a plausible candidate for
orbital decay.  The planet has radius $1.37 \pm 0.04\ R_{\rm Jup}$,
and mass $1.29 \pm 0.10\ M_{\rm Jup}$
\citep{southworth_homogeneous_2011}.  It has an orbital period of 1.34
days, orbits its host star at a distance of $a/R_\star = 5.41$, and
has been frequently observed over the past decade
\citep{wilson_wasp-4b_2008,huitson_gemini_2017}.  The orbit is
consistent with circular (but see \S~\ref{sec:apsidal_precession}),
and has a sky-projected obliquity consistent with zero
\citep{triaud_spin-orbit_2010,beerer_secondary_2011,sanchis-ojeda_starspots_2011}.
The host is a G7V dwarf, with $R_\star = 0.92 \pm 0.13\ R_\odot$,
$M_\star = 0.92 \pm 0.07\ M_\odot$, $T_{\rm eff} = 5400 \pm 90\ {\rm
K}$, and an isochronal age of roughly $7\ {\rm Gyr}$
\citep{southworth_homogeneous_2011,petrucci_no_2013,doyle_accurate_2013}.

We measure new transit times for WASP-4b using data from TESS
(\S~\ref{sec:transits}).  We then merge these times with previous
observations and compare three models for the times
(\S~\ref{sec:timing}): a constant period, a decaying period, and a
slightly eccentric orbit undergoing precession.  We rule out a
constant period, but cannot firmly distinguish between the latter two
possibilities.  Each would have interesting implications
(\S~\ref{sec:implications}).  We rule out the possibility that a
systematic time system offset causes the observed variation
(Appendix~\ref{sec:verify_tess}), and conclude by advocating for
further monitoring of the system (\S~\ref{sec:future}).

\section{New transit times}
\label{sec:transits}

\begin{figure}[t]
    \begin{center}
        \includegraphics[width=0.48\textwidth]{f1.png}
    \end{center}
    \vspace{-0.5cm}
    \caption{
        {\bf TESS transits of WASP-4b.} On the left, black points are
        TESS flux measurements, with a vertical offset applied. Blue
        curves are best-fit models. Rounded midtimes are printed in
        BTJD next to the appropriate transits.  The residuals are
        shown on the right.
        \label{fig:lightcurves}
    }
\end{figure}

\subsection{Observations}
WASP-4 was observed in TESS camera 2 from August 23, 2018 to September
20, during the spacecraft's second sector of science operations.  The
star is designated by the TESS Input Catalog as TIC 402026209
\citep{stassun_TIC_2018}.

@@

\subsection{Measuring the transit times}

\begin{figure}[t]
    \begin{center}
        \leavevmode
        \includegraphics[width=0.48\textwidth]{f2.pdf}
    \end{center}
    \vspace{-0.5cm}
    \caption{
      %TODO: final number
        {\bf TESS saw WASP-4b transit earlier than expected.}
        Points are differences between observed transit times and the
        constant-period ephemeris we fit to literature transits.
        Black points are the more precise half of literatures times;
        gray points are the less precise half.  The blue curves
        represent the $\pm 1\sigma$ credible interval around the
        constant-period ephemeris.  The binned TESS observation (red
        point) arrived $77.8 \pm 10.7\ {\rm seconds}$ before it was
        expected.
        \label{fig:arrived_early}
    }
\end{figure}

Using the times, relative fluxes, and errors determined above, we ran
Box Least Squares (BLS) to estimate the transit epoch, period, and
duration \citep{kovacs_box-fitting_2002}.  Using these parameters, we
isolated each transit to within $\pm$10 transit durations of its
midtime.  In each transit window, we then simultaneously fit for a model
transit, plus a local linear trend.  Our transit model uses the analytic
formulae calculated by \citet{mandel_analytic_2002} and implemented by
\citet{kreidberg_batman_2015}.

We allowed four free parameters per transit: the
time of mid-transit $t_{\rm tra}$, the planet-to-star radius ratio
$R_{\rm p}/R_\star$, and the slope and intercept of the line.
% Our priors were wide uniform distributions centered on the BLS
% estimates for $t_{\rm tra}$ and $R_{\rm p}/R_\star$, and on a slope of
% zero and an intercept of unity for the line.
We fixed the remaining transit parameters as follows.  We set the
eccentricity to zero, and the longitude of periastron to $\pi/2$.  We
assumed a quadratic limb-darkening law, with coefficients interpolated
from the \citet{claret_limb_2017} tables.  For the period, $a/R_\star$,
and inclination, we adopted the values from~\citet{petrucci_no_2013}.
These are in reasonable agreement with the parameters reported by
\citet{gillon_improved_2009}, \citet{southworth_high-precision_2009},
and \citet{huitson_gemini_2017}.

Having defined the free and fixed parameters, we calculated an initial
guess for the free parameters by maximizing the likelihood.  We then
sampled over the posterior using the algorithm proposed by
\citet{goodman_ensemble_2010} and implemented by
\citet{foreman-mackey_emcee_2013}.  During this process, we used the
\texttt{corner} software to check that our posteriors were being
well-sampled~\citep{corner_2016}.

After performing the initial fit, we set the photometric error bars to
be equal to the standard deviation of the in-transit points of the
residual lightcurve.  We then reperformed the same fitting procedure,
using the empirically determined photometric errors.

% TODO: final numbers
To verify that the measured uncertainties are plausible, we computed the
reduced $\chi^2$ for a linear ephemeris fit to the measured TESS transit
times.  We found that $\chi^2 = 8.4$, with 16 degrees of freedom.
Visually inspecting residuals showed that the error variance was being
overestimated, so we multiplied the measured TESS errors by a factor
$f=0.73$, forcing a reduced $\chi^2$ of unity.  This lowered the mean
midtime uncertainty for each transit from 30 to 22
seconds.

Figure~\ref{fig:lightcurves} shows the light curves, best-fit transit
models, and residuals.  Table~1 reports the mid-transit times and
their uncertainties.
Taking the mean and standard deviation of the measured planet to star
radius ratio, we find $R_{\rm p}/R_\star=0.1538\pm0.0013$, in
agreement with previous determinations
\citep{wilson_wasp-4b_2008,gillon_improved_2009,winn_transit_2009,southworth_high-precision_2009}.
Binning the residuals to 1-hour windows and taking the median absolute
deviation, the residual lightcurves have a MAD of $594\,{\rm
ppm}\,{\rm hr}^{1/2}$.  WASP-4 has a TESS-band magnitude of $T=11.78$
\citep{stassun_TIC_2018}, corresponding to a predicted photon-counting
noise\footnote{\url{github.com/lgbouma/tnm}} of $410\,{\rm ppm}\,{\rm
hr}^{1/2}$.  At $T=11.78$, with a 9-pixel aperture read noise is
expected to contribute an additional $202\,{\rm ppm}\,{\rm hr}^{1/2}$ in
quadrature.  The predicted zodiacal background contribution is
$673\,{\rm ppm}\,{\rm hr}^{1/2}$, based on the
\citet{winn_photonflux_2013} model, which \citet{Sullivan_2015}
used in planet detection simulations.  The predicted level
of zodiacal noise seems to have been an overestimate, given the observed
RMS of WASP-4.  A more detailed assessment of TESS's photometric
performance is beyond the scope of this work.

@@

\begin{figure*}[t]
    \begin{center}
        \leavevmode
        \includegraphics[width=0.9\textwidth]{f3.pdf}
    \end{center}
    \vspace{-0.5cm}
    \caption{
        {\bf Timing residuals and best-fit models for WASP-4b.}
        Points are differences between observed transit times and a
        linear ephemeris fit to all the times.  The constant period
        model is a poor fit to the observed times (gray line).  Black
        points are the more precise half of measured times; gray points
        are the less precise half.  The blue curve is the best-fit model
        with a constant period derivative.  The orange curve is the
        best-fit model assuming apsidal precession causes the observed
        timing deviations.
        \label{fig:times}
    }
\end{figure*}

Table~1 shows the transit times used in our analysis.  We included
data from peer-reviewed literature for which the analysis was based on
observations of a single, complete transit, and for which the midpoint
was fit as a free parameter. We also required that the time system be
clearly documented.

Table~2 shows the occultation times.  Since there are fewer available,
we included all the occultation measurements we could find. The
tabulated values have been corrected for the light-travel time across
the diameter of the orbit by subtracting $2a/c = 22.8$ seconds from
the observed time.

>>>>

\documentclass[12pt,twocolumn,tighten]{aastex62}
%\pdfoutput=1 %for arXiv submission
\usepackage{amsmath,amstext,amssymb}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage[figure,figure*]{hypcap}
\usepackage{graphics,graphicx}
\usepackage{hyperref}
\usepackage{comment}
\usepackage{todonotes}

@@

\shortauthors{Bouma et al.}
\shorttitle{Early Arrival of WASP-4\lowercase{b}}

\NewPageAfterKeywords

@@

\title{The Early Arrival of WASP-4\lowercase{b}}

@@

\author[0000-0002-0514-5538]{L. G. Bouma}
\affiliation{ Department of Astrophysical Sciences, Princeton
University, 4 Ivy Lane, Princeton, NJ 08540, USA}
%
\author[0000-0002-4265-047X]{J. N. Winn}
\affiliation{ Department of Astrophysical Sciences, Princeton
University, 4 Ivy Lane, Princeton, NJ 08540, USA}
%
\author[0000-0002-0875-8401]{J.-M. D\'esert}
\affiliation{ API, University of Amsterdam, P.O. Box 94249, 1090 GE
Amsterdam, The Netherlands}
%
\author[0000-0002-3481-9052]{K. G. Stassun}
\affiliation{Vanderbilt University, Department of Physics \& Astronomy,
6301 Stevenson Center Lane, Nashville, TN 37235, USA}
\affiliation{Fisk University, Department of Physics, 1000 17th Avenue
N., Nashville, TN 37208, USA}
%
\author[0000-0002-6939-9211]{T. Daylan}
\affiliation{ Department of Physics and Kavli Institute for Astrophysics
and Space Research, Massachusetts Institute of Technology, Cambridge, MA
02139, USA }
%
\author[0000-0002-7084-0529]{S. Kane}
\affiliation{Department of Earth Sciences, University of California,
Riverside, CA 92521, USA}
% 
%-------------------------------------
% TESS Mission Architects:
% These authors should be listed in this order
% see https://spacebook.mit.edu/pages/viewpage.action?pageId=24543276
%-------------------------------------
%
\author{G. R. Ricker} % grr@space.mit.edu WAITING
\affiliation{ Department of Physics and Kavli Institute for Astrophysics
and Space Research, Massachusetts Institute of Technology, Cambridge, MA
02139, USA }
%
\author[0000-0001-6763-6562]{R. Vanderspek} % roland@space.mit.edu WAITING
\affiliation{ Department of Physics and Kavli Institute for Astrophysics
and Space Research, Massachusetts Institute of Technology, Cambridge, MA
02139, USA }
%
\author[0000-0001-9911-7388]{D. W.~Latham} % dlatham@cfa.harvard.edu WAITING
\affiliation{ Harvard-Smithsonian Center for Astrophysics, 60 Garden
Street, Cambridge, MA 02138, USA }
%
\author{S. Seager} % seager@mit.edu WAITING
\affiliation{ Department of Earth, Atmospheric, and Planetary Sciences,
Massachusetts Institute of Technology, Cambridge, MA 02139, USA }
%
\author[0000-0002-4715-9460]{J. M.~Jenkins} % jon.jenkins@nasa.gov WAITING
\affiliation{ NASA Ames Research Center, Moffett Field, CA 94035, USA }
%
%-------------------------------------
% 3 representatives of each of SPOC, POC, TSO, for a total of 9. 
%These 9 authors should be listed in alphabetical order
%-------------------------------------
% 3 TSO COAUTHORS
%
\author[0000-0002-3321-4924]{Z. Berta-Thompson} % Zachory.BertaThompson@Colorado.EDU WAITING
\affiliation{Department of Astrophysical and Planetary Sciences,
University of Colorado, Boulder, CO 80309, USA}
%
\author[0000-0001-8812-0565]{J. Rodriguez} % rodriguez.jr.joey@gmail.com WAITING
\affiliation{Harvard-Smithsonian Center for Astrophysics, 60 Garden
Street, Cambridge, MA 02138, USA }
%
\author[0000-0001-8020-7121]{K. Col\'on} % knicole.colon@nasa.gov WAITING
\affiliation{NASA Goddard Space Flight Center, Exoplanets and Stellar
Astrophysics Laboratory (Code 667), Greenbelt, MD 20771, USA}
% 3 SPOC COAUTHORS
%
\author[0000-0002-6778-7552]{J. D. Twicken} % Joseph.Twicken@nasa.gov WAITING
\affiliation{ NASA Ames Research Center, Moffett Field, CA 94035, USA }
\affiliation{ SETI Institute, Mountain View, CA 94043, USA}
%
\author{B. Wohler} % bill.wohler@nasa.gov WAITING
\affiliation{ NASA Ames Research Center, Moffett Field, CA 94035, USA }
\affiliation{ SETI Institute, Mountain View, CA 94043, USA}
%
\author[0000-0003-1963-9616]{D. A. Caldwell} % douglas.Caldwell@nasa.gov % WAITING
\affiliation{ NASA Ames Research Center, Moffett Field, CA 94035, USA }
\affiliation{ SETI Institute, Mountain View, CA 94043, USA}
%
%\\ \textcolor{red}{3 POC coauthors:}
\author{POC1}
\author{POC2}
\author{POC3}

\begin{abstract}
  %TODO: final numbers
  Combining data from the Transiting Exoplanet Survey Satellite (TESS)
  with previous studies, we show that the transit times of WASP-4b are
  incompatible with a constant orbital period.  In particular, the
  transits seem to have arrived $77.8 \pm 10.8$ seconds early, and the
  period appears to be shrinking by $\dot{P}=-12.1 \pm 1.2$
  milliseconds per year.   From TESS observations of WASPs-6b, -18b,
  and -46b, we show that a systematic offset between the TESS times
  and the barycentric reference sufficient to explain WASP-4b is ruled
  out at 6.3$\sigma$.  The timing variations seem to be astrophysical,
  and could be caused by apsidal precession, or perhaps by tidal
  decay.  The latter seems implausible on theoretical grounds, as it
  would require a stellar quality factor $Q_\star' \approx
  3\times10^4$.  Apsidal precession would imply an eccentricity
  $e\approx1.6\times 10^{-3}$, and the possibility of measuring the
  planet's Love number.  A Doppler shift from WASP-4's acceleration
  towards us could account for at most one quarter of the observed
  period change (at $2\sigma$).  Further observations will help
  confirm the reality of the timing variation, and eventually
  determine its cause.
\end{abstract}

@@

Long-term monitoring of hot Jupiter transit and occultation times
should eventually reveal two distinct processes.  First, whenever the
summed spin and orbital angular momenta in a hot Jupiter system fall
below a critical value, the system becomes unstable to tidal decay
\citep{counselman_outcomes_1973,hut_stability_1980}.  Most hot
Jupiters satisfy this condition; their orbits should be shrinking
\citep{levrard_falling_2009,matsumura_tidal_2010}.  The rate of decay
is uncertain, and depends on how friction dissipates energy carried by
gravitational tides (\citealt{ogilvie_tidal_2014} gives a review).  If
we could directly measure orbital decay rates, it would inform our
understanding of tidal dissipation.  This is important because the
dissipation rates determine the fate of the system.  During its
inspiral, if the planet fills its Roche lobe below the stellar
surface, the drag force from the stellar envelope accelerates the
inspiral, leading to a `direct-impact' merger
\citep{metzger_optical_2012,macleod_planetary_2018} Alternatively, if
the planet fills its Roche lobe above above the stellar surface, then
it could either be disrupted into an accretion disk or it could slowly
transfer mass to the star
\citep{metzger_optical_2012,valsecchi_tidally-driven_2015,jackson_tidal_2016}.
Further afield, the outcome of the eventual interaction between the
Earth and red giant Sun also hinges on the efficiency of tidal
dissipation \citep{rasio_tidal_1996}.

@@

This work focuses on the hot Jupiter WASP-4b, which orbits its host
star every 1.34 days, at a distance of $a/R_\star = 5.46$
\citep{wilson_wasp-4b_2008,hoyer_tramos_2013}.  WASP-4b is a plausible
candidate for orbital decay, in part because it has been frequently
observed over the past decade.  The host is a G7V dwarf, and is on the
main sequence (\S~\ref{sec:system_parameters}).  The orbit is
consistent with being circular (but see
\S~\ref{sec:apsidal_precession}), and having sky-projected obliquity
of zero
\citep{triaud_spin-orbit_2010,beerer_secondary_2011,sanchis-ojeda_starspots_2011}.

We measure new transit times for WASP-4b using TESS lightcurves
(\S~\ref{sec:observations},~\ref{sec:measurement}), and calculate the
system's parameters empirically (\S~\ref{sec:system_parameters}).  We
merge the new times with previous observations and compare three
models (\S~\ref{sec:timing}): a constant period, a decaying period,
and a slightly eccentric orbit undergoing precession.  We rule out a
constant period, but cannot firmly distinguish between the latter two
possibilities.  Each would have interesting implications
(\S~\ref{sec:implications}).  We rule out the possibility that a
systematic time system offset causes the observed variation
(Appendix~\ref{sec:verify_tess}), and conclude by advocating for
further monitoring of the system (\S~\ref{sec:future}).

\section{New transits and system parameters}
\begin{figure}[t]
    \begin{center}
        \includegraphics[width=0.48\textwidth]{f1.png}
    \end{center}
    \vspace{-0.5cm}
    \caption{
        {\bf TESS transits of WASP-4b.} On the left, black points are
        TESS flux measurements, with a vertical offset applied. Blue
        curves are best-fit models. Rounded midtimes are printed in
        BTJD next to the appropriate transits.  The residuals are
        shown on the right.
        \label{fig:lightcurves}
    }
\end{figure}

\subsection{Observations}
\label{sec:observations}

WASP-4 was observed in TESS camera 2 from August 23, 2018 to September
20, during the spacecraft's second sector of science operations.  The
star is designated by the TESS Input Catalog as TIC 402026209
\citep{stassun_TIC_2018}.

@@

\subsection{Measuring the transit times}
\label{sec:measurement}

\begin{figure}[t]
    \begin{center}
        \leavevmode
        \includegraphics[width=0.48\textwidth]{f2.pdf}
    \end{center}
    \vspace{-0.5cm}
    \caption{
      %TODO: final number
        {\bf TESS saw WASP-4b transit earlier than expected.}
        Points are differences between observed transit times and the
        constant-period ephemeris we fit to literature times.  Black
        points are the more precise half of literature times; gray
        points are the less precise half.  Blue curves represent the
        $\pm 1\sigma$ credible interval around the constant-period
        ephemeris.  The binned TESS observation (red point) arrived
        $77.8 \pm 10.8\ {\rm seconds}$ before it was expected.
        \label{fig:arrived_early}
    }
\end{figure}

Using the cleaned PDC lightcurve, we ran Box Least Squares (BLS) to
estimate the transit epoch, period, and duration
\citep{kovacs_box-fitting_2002}.  Using these parameters, we isolated
each transit to within $\pm$10 transit durations of its midtime.  In
each transit window, we then simultaneously fit for a model transit,
plus a local linear trend.  Our transit model uses the analytic
formulae calculated by \citet{mandel_analytic_2002} and implemented by
\citet{kreidberg_batman_2015}.

We allowed four free parameters per transit: the time of mid-transit
$t_{\rm tra}$, the planet-to-star radius ratio $R_{\rm p}/R_\star$,
and the slope and intercept of the line.
% Our priors were wide uniform distributions centered on the BLS
% estimates for $t_{\rm tra}$ and $R_{\rm p}/R_\star$, and on a slope
% of zero and an intercept of unity for the line.
We fixed the remaining transit parameters as follows.  We set the
eccentricity to zero, and the longitude of periastron to $\pi/2$.  We
assumed a quadratic limb-darkening law, with coefficients interpolated
from the \citet{claret_limb_2017} tables.  For the period,
$a/R_\star$, and inclination, we adopted the values
from~\citet{petrucci_no_2013}.  These are in reasonable agreement with
the parameters reported by \citet{gillon_improved_2009},
\citet{southworth_high-precision_2009}, and
\citet{huitson_gemini_2017}.

Having defined the free and fixed parameters, we calculated an initial
guess for the free parameters by maximizing the likelihood.  We then
sampled over the posterior using the algorithm proposed by
\citet{goodman_ensemble_2010} and implemented by
\citet{foreman-mackey_emcee_2013}.
% During this process, we used the \texttt{corner} software to check
% that our posteriors were being well-sampled~\citep{corner_2016}.
After completing an initial fit, we set the photometric error bars to
be equal to the standard deviation of the in-transit points of the
residual lightcurve.  We then reperformed the fitting procedure, using
the empirically determined photometric errors.

% TODO: final numbers
To verify that the measured uncertainties are plausible, we computed
the reduced $\chi^2$ for a linear ephemeris fit to the measured TESS
transit times.  We found that $\chi^2 = 8.4$, with 16 degrees of
freedom.  Visually inspecting residuals showed that the error variance
had been overestimated, so we multiplied the measured TESS errors by a
factor $f=0.73$, forcing a reduced $\chi^2$ of unity.  This lowered
the mean midtime uncertainty for each transit from 30 to 22 seconds.

Figure~\ref{fig:lightcurves} shows the light curves, best-fit models,
and residuals.  Table~1 reports the mid-transit times and their
uncertainties.  Taking the mean and standard deviation of the measured
planet to star radius ratio, we find $R_{\rm
p}/R_\star=0.1538\pm0.0013$, in agreement with previous determinations
\citep{wilson_wasp-4b_2008,gillon_improved_2009,winn_transit_2009,southworth_high-precision_2009}.
Binning the residuals to 1-hour windows and taking the median absolute
deviation, the lightcurves have a MAD of $594\,{\rm ppm}\,{\rm
hr}^{1/2}$.  The predicted\footnote{\url{github.com/lgbouma/tnm}}
photon-counting noise is $410\,{\rm ppm}\,{\rm hr}^{1/2}$.  At
$T\approx11.8$, with a 9-pixel aperture read noise is expected to
contribute an additional $202\,{\rm ppm}\,{\rm hr}^{1/2}$ in
quadrature.  The predicted zodiacal background contribution is
$673\,{\rm ppm}\,{\rm hr}^{1/2}$, based on the
\citet{winn_photonflux_2013} model, which \citet{Sullivan_2015} used
in planet detection simulations.  This zodiacal noise prediction seems
to have been an overestimate, given the observed MAD of WASP-4.  A
more detailed assessment of TESS's photometric performance is beyond
the scope of this work.

\subsection{Star and planet parameters}
\label{sec:system_parameters}

\begin{deluxetable}{lccc}
\tabletypesize{\scriptsize}
\tablecaption{Selected system parameters of WASP-4b\label{tbl:params}}
\tablenum{2}

\tablehead{
\colhead{Parameter} & \colhead{Value} & \colhead{Error} & \colhead{Comment}
}

\startdata
{\it Transit/RV parameters:} & & & \\
  $R_{\rm p}/R_\star$                        & $0.1538$               & $\pm 0.0013$                & A \\
  $i$~[deg]                                  & $88.52$                & $+0.39$, $-0.26$            & B \\
  $a/R_\star$                                & $5.463$                & $+0.025$, $-0.020$          & B \\
  $K$~[m~s$^{-1}$]                           & $241.1$                & $+2.8$, $-3.1$              & C \\
{\it Stellar parameters:} & & & \\
  $T_{\rm eff}$~[K]                          & $5400$                 & $\pm 90$                    & D \\
  $\log g_\star$~[cgs]                       & $4.47$                 & $\pm 0.11$                  & D \\
  $[{\rm Fe/H}]$                             & $-0.07$                & $\pm 0.19$                  & D \\
  $F_{\rm bol}$~[erg~cm$^{-2}$~s$^{-1}$]     & $2.802\times10^{-10}$  & $\pm 0.076\times10^{-10}$   & E \\
  $\pi$~[mas]                                & $3.7145$               & $0.0517$                    & G \\
  $R_\star$~[R$_{\odot}$]                    & $0.893$                & $\pm 0.034$                 & F \\
  $\rho_\star$~[g~cm$^{-3}$]                 & $1.722$                & $+0.024$, $-0.019$          & A \\
  $M_\star$~[M$_{\odot}$]                    & $0.870$                & $+0.085$, $-0.077$          & F \\
  $T$ magnitude                              & $11.778$               & $\pm 0.018$                 & H \\
{\it Planetary parameters:} & & & \\
  $a$~[AU]                                   & $0.0227$               & $\pm 0.0007$                & F \\
  $M_p$~[M$_{\rm Jup}$]                      & $1.192$                & $+0.091$, $-0.086$          & F \\
  $R_p$~[R$_{\rm Jup}$]                      & $1.337$                & $\pm 0.048$                 & F \\
\enddata

\tablecomments{
  (A) From TESS lightcurves (\S~\ref{sec:measurement}).
  (B) \citet{hoyer_tramos_2013}.
  (C) \citet{triaud_spin-orbit_2010}.
  (D) From HARPS spectra \citep{doyle_accurate_2013}.
  (E) \citet{stassun_accurate_2017}.
  (F) From functions of stellar, transit or RV parameters as discussed in text.
  (G) Gaia DR2.
  (H) \citet{stassun_TIC_2018}.
}

\end{deluxetable}

\citet{doyle_accurate_2013} found accurate spectroscopic parameters
for WASP-4 using high S/N HARPS spectra.  Using their (T$_{\rm eff}$,
$\log g$, $[{\rm Fe/H}]$) and the available {\it Gaia} parallaxes, we
compute the remaining system parameters empirically.

\citet{stassun_accurate_2017} found the bolometric flux from
broad-band photometry.  Combining the implied angular radius with the
{\it Gaia} DR2 parallax, and adjusting for the parallax offset
\citep{stassun_evidence_2018}, we obtain the stellar radius.  For the
stellar mass, we started with the $a/R_\star$ value found by
\citet{hoyer_tramos_2013} from 26 ground-based transits.  The implied
semi-major axis and stellar density follow \citep{seager_unique_2003}.
Combining the density and radius gives the stellar mass.  We get the
planet radius using our measured $R_{\rm p}/R_\star$.  Finally, for
the planet mass we use the observed RV semi-amplitude and inclination
angle from the transit
\citep{triaud_spin-orbit_2010,hoyer_tramos_2013}.

The resulting parameters are in Table~3, and broadly agree with those
of previous authors
\citep{wilson_wasp-4b_2008,gillon_discovery_2009,winn_transit_2009,southworth_homogeneous_2011,petrucci_no_2013}.
We use them for the remaining analysis.  Comparing the empirical
result to the Yonsei-Yale evolutionary track predicted by the
spectroscopic parameters, we see that WASP-4  has an age $\approx
7\,{\rm Gyr}$, and is securely in the main-sequence stage of
evolution.

@@

\begin{figure*}[t]
    \begin{center}
        \leavevmode
        \includegraphics[width=0.9\textwidth]{f3.pdf}
    \end{center}
    \vspace{-0.5cm}
    \caption{
        {\bf Timing residuals and best-fit models for WASP-4b.}
        Points are differences between observed transit times and a
        linear ephemeris fit to all the times.  The constant period
        model is a poor fit to the observed times (gray line).  Black
        points are the more precise half of measured times; gray
        points are the less precise half.  The blue curve is the
        best-fit model with a constant period derivative.  The orange
        curve is the best-fit model assuming apsidal precession causes
        the observed timing deviations. The binned TESS point (star)
        is for display purposes only.
        \label{fig:times}
    }
\end{figure*}

Table~1 shows the transit times used in our analysis.  We included
data from peer-reviewed literature for which the analysis was based on
observations of a single, complete transit, and for which the midpoint
was fit as a free parameter. We also required that the time system be
clearly documented.  The occultation times are in Table~3.  Since
there are fewer available, we included all the occultation
measurements we could find. The tabulated values have been corrected
for the light-travel time across the diameter of the orbit by
subtracting $2a/c = 22.8$ seconds from the observed time.

We used the homogeneous \citet{hoyer_tramos_2013} timing study as our
starting point.  Though their latest two epochs fell slightly below
the prediction of a linear ephemeris (Panel C of their Figure~6), they
did not find a need to fit a quadratic function to the O-C values.  We
verified that all times were on the BJD$_{\rm TDB}$ time standard
using the \citet{eastman_achieving_2010} calculator.  We set the
zero-point epoch to be as close as possible to the average of the
midtimes, weighted as the inverse midtime variance. This minimizes the
covariance between the transit epoch and the period
\citep{gibson_gemini_2013}.
